---
title: 'Lesson3: Advanced Geoms and Statistics'
author: "Tania Moulik"
date: "June 13, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description
This code contain Exercies and Activities for Lesson 3. This is a R Markdown
file. To Run it, click on "Run" and choose "Run all chunks below". The plots
are visible under the individual chunks below.

This code can be converted to a "PDF" by clicking on "Knit" and selecting
"PDF"

### Libraries used:
Lock5Data - contains some pre-made datasets. 
tibble
dplyr


Prior to use, install the following packages:
install.packages("ggplot2")
install.packages("tibble")
install.packages("dplyr")
install.packages("gridExtra")
install.packages("Lock5Data")
install.packages("maps")
install.packages("mapproj")
install.packages("corrplot")

Used datafiles and sources:
a) gapminder.csv - Modified dataset from various datasets available at:
https://www.gapminder.org/data/
b) xAPI-Edu-Data.csv:
https://www.kaggle.com/aljarah/xAPI-Edu-Data/data
c) LoanStats.csv:
Loan Data from Lending Tree - https://www.lendingclub.com/info/download-data.action
d) Lock5Data




```{r load}
require("ggplot2")
require("tibble")
require("dplyr")
require("Lock5Data")
require("zoo")
require("corrplot")
require("maps")
require("mapproj")

```
Set current working directory to :Beginning-Data-Visualization-with-ggplot2-and-R

```{r read}
#Load the data files
df <- read.csv("../data/gapminder-data.csv")
df2 <- read.csv("../data/xAPI-Edu-Data.csv")
df3 <- read.csv("../data/LoanStats.csv")

```


###Exercise 
Create a bubble chart

```{r bubble}
dfs <- subset(df,Country %in% c("Germany","India","China","United States","Japan"))

ggplot(dfs,aes(x=Year,y=Electricity_consumption_per_capita)) + geom_point(aes(size=population,color=Country))+
  coord_cartesian(xlim=c(1950,2020))+
labs(subtitle="Electricity consumption vs Year",
       title="Bubble chart")+ylab("Electricity consumption")+
  scale_size(breaks=c(0,1e+8,0.3e+9,0.5e+9,1e+9,1.5e+9),range=c(1,5))

```

### Exercise: Creating density plots

```{r density}
#df3

#Let's do a histogram first and sub divide into the different grades.
ggplot(df3,aes(x=loan_amnt)) + geom_histogram() + facet_wrap(~grade)

# We cannot see the shapes of the E,F,G grades very clearly. Also all the
#grades have different histogram counts. It would be better to use a density
#plot to compare.


ggplot(df3,aes(x=loan_amnt)) + geom_density() + facet_wrap(~grade)


```

###Superimposing plots 

```{r superimpose}

ggplot(df3,aes(x=loan_amnt)) + geom_density(aes(fill=grade),alpha=1/2) +
  scale_fill_brewer(palette="Dark2") + xlab("Loan Amount") + theme_light()

```

###Activity A
```{r activity-A}

ggplot(RestaurantTips,aes(x=Tip)) + geom_density(aes(color=Day),alpha=0.2) +
  scale_fill_brewer(palette="Dark2") + xlab("Tip") + scale_x_continuous(breaks=seq(0,15,1)) + theme_classic()

```

### Exercise: To create a time series plot
```{r time-series}

#Read the data
df_fb <- read.csv("../data/FB.csv")

#Glimpse
glimpse(df_fb)
```

```{r}
#Convert date to Character
df_fb$Date <- as.Date(df_fb$Date)
```
```{r ts-plot}
#Daily data plot
ggplot(df_fb,aes(x=Date,y=Close, group=1)) + 
  geom_line(color="black",na.rm=TRUE)+
  ggtitle("Daily Closing Stock Prices: Facebook") + 
  theme(plot.title = element_text(lineheight=.7, face="bold")) + 
  scale_x_date(date_breaks='3 month')+
  theme_bw()
```


### Activity B

Plot the mean monthly closing stock prices.

```{r ts-plot-sum}
#Use strftime to get the month for each date
df_fb$Month <- strftime(df_fb$Date,"%m")
df_fb$Month <- as.numeric(df_fb$Month)
ggplot(df_fb, aes(Month,Close)) + 
  geom_point(color="red",alpha=1/2,position = position_jitter(h=0.0,w=0.0))+
  geom_line(stat='summary',fun.y=mean, color="blue",size=1)+
  scale_x_continuous(breaks=seq(0,13,1))+
  ggtitle("Monthly Closing Stock Prices: Facebook")+theme_classic()

```

## Maps

```{r maps}
states_map <- map_data("state")
```

```{r}
glimpse(states_map)
```

The map_data() function returns a data frame with the following columns: 

long - Longitude

lat - Latitude

group - This is a grouping variable for each polygon

A region or subregion might have multiple polygons, for example, if it includes islands.

```{r mapplot}
ggplot(states_map, aes(x=long, y=lat, group=group)) +    geom_polygon(fill="white", colour="black")

ggplot(states_map, aes(x=long, y=lat, group=group)) +    
  geom_path() + coord_map("mercator")

```


### World map data

```{r world}
# Get map data for world 
world_map <- map_data("world") 
#world_map
``` 

### Map of Europe
```{r europe}

europe <- map_data("world", region=c("Germany", "Spain", "Italy",                                        "France","UK","Ireland")) 
ggplot(europe, aes(x=long, y=lat, group=group, fill=region)) +    geom_polygon(colour="black") + scale_fill_brewer(palette="Set3")


```
### Activity C

Create a map with regions that are colored according to variable values

```{r activity-maps}
USStates$Statelower <- as.character(tolower(USStates$State))
glimpse(USStates)

us_data <- merge(USStates,states_map,by.x="Statelower",by.y="region")
head(us_data)

```
Voter Chart for 2012 Elections

```{r}
ggplot(us_data, aes(x=long, y=lat, group=group, fill=ObamaVote)) +    geom_polygon(colour="black") +
  coord_map("mercator")+scale_fill_gradient(low="red",high="blue")

```
Compare the above plot to the voting chart to see which states voted 
primarily for Obama in the 2012 elections in which Obama won. 

Electoral College Maps: https://www.nytimes.com/2016/08/23/upshot/50-years-of-electoral-college-maps-how-the-us-turned-red-and-blue.html

## Trends, correlations and statistical summaries

### Statistical summaries

Let's do the same time series plot as before, but now along with the mean we
would also like to plot the median and the 10% and 90% quantiles.

```{r ts-sum}
#Using stat-summary - Plot Mean closing price per month.
df_fb$Month <- as.numeric(df_fb$Month)
ggplot(df_fb, aes(Month,Close)) + 
  geom_point(color="red",alpha=1/2,position = position_jitter(h=0.0,w=0.0))+
  geom_line(stat='summary',fun.y=mean, color="blue",size=1)+
  geom_line(stat='summary',fun.y=median, color="orange",size=1)+
    geom_line(stat='summary',fun.y=quantile, fun.args=list(probs=0.1),
            linetype=2,color='green',size=1.)+
  geom_line(stat='summary',fun.y=quantile, fun.args=list(probs=0.9),
            linetype=2,color='green',size=1.)+
  scale_x_continuous(breaks=seq(0,13,1))+
  ggtitle("Monthly Closing Stock Prices: Facebook")+theme_classic()

```

### Trends and Scatterplots
One can study trends in data by looking at scatter plots between two variables. This reveals if one variable is related to another variable. 

```{r scatplot}
ggplot(dfs, aes(gdp_per_capita,Electricity_consumption_per_capita)) + geom_point(aes(color=Country))+xlim(0,30000)+ stat_smooth(method=lm)

ggplot(dfs, aes(gdp_per_capita,Electricity_consumption_per_capita,color=Country)) + geom_point() + stat_smooth(method=lm)
  
```


### Correlation plot
```{r corr}
#Use only continuous variables columns. Drop "Year","Country"
dfs1 <- dfs[,colnames(dfs)[4:9]]
#Remove NA's or correlation won't work
dfs1 <- na.omit(dfs1)
M <- cor(dfs1)
corrplot(M,method="circle")
```

The plot looks messy because of the long names. Let's change the names to
shorter names.

```{r corother}
colnames(dfs1) <- c("gdp","electricity","mort","pov","bmi_m","bmi_f")
M <- cor(dfs1)
corrplot(M,method="number")
```


One can see that the positively correlated variables (in Blue and Darkblue) are:

1. GDP and Electricity consumption

2. Electricity consumption and BMI's

3. BMI's of males and females

4. Poverty and under5mortality

The one's which have negative corelation are:

1. mortality and electricity consumption

2. GDP and Poverty.

3. GDP and Mortality

4. Poverty and BMI

One can also try other methods for the correlation plots:

```{r corr-other}
corrplot(M,method="pie")

corrplot(M,method="ellipse")
```

In the first plot the fraction of the pie, gives an idea of how strong the correlation is and the color gives whether its positively or negatively correlated. In the second case, the width of the ellipse gives the indication 
of the correlation. 

a) The "number" option gives us an exact number and there is no ambiguity. However, when one is presenting to an audience, it makes it difficult to
read.

b) The "circle" option is great visually as its color coded. But the 'ellipse'
or 'pie' option is even better because one gets a feel for the size as
well as type of correlation.

### Activity D

Studying Correlations and making a scatter plot.
```{r activityd}

t <- subset(df3,grade=="A")
ggplot(t, aes(total_pymnt_inv,total_rec_prncp,color=grade)) + geom_point() + stat_smooth(method=lm)

ggplot(t, aes(funded_amnt,total_pymnt_inv,color=grade)) + geom_point() +
  stat_smooth(method=lm,color=2)

```
